---
title: "DATA 606 Data Project Proposal"
authors: "Fomba Kassoh & Souleymane Doumbia"
date: "2023-10-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Data Preparation


```{r}
library(dplyr)
library(readr)
library(stringr)
library(tidyverse)
```


```{r}
# load data
covid_data <- read_csv("https://raw.githubusercontent.com/hawa1983/DATA606-Project/main/covid_data1.csv")
glimpse(covid_data)
```


### Research question 

**You should phrase your research question in a way that matches up with the scope of inference your dataset allows for.**

Among individuals tested for a specific condition in October 2020, how does age group, sex, and race influence the likelihood of requiring hospitalization? Specifically, are certain age groups, genders, or racial/ethnic groups at a higher risk of severe outcomes as indicated by hospitalization?

### Cases 

**What are the cases, and how many are there?**
Each case represents a Covid-19 patient in the united states. There 203,224 observations in the given data set.


### Data collection 

**Describe the method of data collection.**
Data is collected by the Center for Disease Control and Prevention (CDC). Data is submitted by hospitals.


### Type of study 

**What type of study is this (observational/experiment)?**
This is an observational study.

### Data Source 

**If you collected the data, state self-collected. If not, provide a citation/link.**
Data is collected by CDC and is available online here: https://data.cdc.gov/Case-Surveillance/COVID-19-Case-Surveillance-Public-Use-Data/vbim-akqf. For this project, data was extracted by downloading the csv format table.



### Dependent Variable

**What is the response variable? Is it quantitative or qualitative?**

The response variable is "Death Status (Did the patient die as a result of covid?)," which is a categorical variable. It is recorded by hospitals that admit covid patients.

### Independent Variable(s)

The independent variables are race and gender


### Relevant summary statistics 

**Provide summary statistics for each the variables. Also include appropriate visualizations related to your research question (e.g. scatter plot, boxplots, etc). This step requires the use of R, hence a code chunk is provided below. Insert more code chunks as needed.**

Covid hospitalization by race
```{r by race}
ggplot(covid_data, aes(race_ethnicity_combined)) +
  geom_bar() +
  labs(title = "Covid Hospitalization by Race",
       x = "Race",
       y = "Frequency") +
  coord_flip()
```

Proportion of Covid infection by gender

```{r prop by sex}
covid_data |> 
  count(sex)|>
  mutate(proportion = round(n/sum(n),6))
```

Proportion pf Icu admission
```{r prop by icu}
covid_data |> 
  count(icu_yn)|>
  mutate(proportion = round(n/sum(n),6))
```

Distribution of Covid infection by age group
```{r prop by icu}
covid_data |> 
  count(age_group)|>
  mutate(proportion = round(n/sum(n),6))
```


```{r by race}
ggplot(covid_data, aes(age_group)) +
  geom_bar() +
  labs(title = "Covid Hospitalization by Race",
       x = "Age Group",
       y = "Frequency") +
  coord_flip()
```

Time series of covid deaths.
```{r}

death_yes <- covid_data %>%
  filter(death_yn == "Yes") %>%
  mutate(cdc_case_earliest_dt = as.Date(cdc_case_earliest_dt, format = "%m/%d/%Y")) %>%
  count(cdc_case_earliest_dt) %>%
  arrange(cdc_case_earliest_dt)

# Assuming you have a data frame named 'mydata'
ggplot(death_yes, aes(x = cdc_case_earliest_dt, y = n)) +
  geom_line() +           # Line plot
  labs(x = "Date", y = "No. Deaths") +  # Labels for axes
  ggtitle("Time Series Plot of Covid Death")   # Title for the plot


```
